---
title: "Project 3: STAT302PACKAGE Tutorial"
author: "David Zeng"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{STAT302PACKAGE Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(STAT302PACKAGE)
```

# Introduction

The STAT302PACKAGE contains data and four main functions (in addition to some 
other minor ones):

* ```my_t.test```, a function which computes a one sample t-test 
* ```my_lm```, a function which fits a linear model 
* ```my_knn_Cv```, a function which trains a n-nearest neighbors model using 
k-fold cross-validation
* ```my_rf_cv``, a function which trains a random forest model using k-fold 
cross-validation

To install and use this package, run the following commands in console:

```{r, eval=FALSE}
# install and load package
devtools::install_github("dzeng8/STAT302PACKAGE")
library(STAT302PACKAGE)
```

# ```my_t.test``` Tutorial

We will demonstrate the ```my_t.test``` function with ```lifeExp``` data from
the ```my_gapminder``` dataset. 

The input data are measurements of people's life expectancy at birth, in years.

### two-sided Test

We run a t-test with the null hypothesis that the mean life expectancy is 60 years,
with the alternative hypothesis being that the life expectancy is not 60 years.

```{r}
# run two-sided t-test with null hypothesis of mean as 60
results <- my_t.test(my_gapminder$lifeExp, "two.sided", 60)
results
```

Using a p-value cut-off at $\alpha = 0.05$, we can see that a p-value of 
`r results[[4]]` implies that we cannot
reject the null hypothesis. In other words, there is no evidence to reject the
null hypothesis that the mean life expectancy is 60 in support of the alternative
hypothesis that the mean life expectancy is not 60.

### one-sided test, less

We run a t-test with the null hypothesis that the mean life expectancy is 60 years,
with the alternative hypothesis being that the life expectancy is less than 60 years.

```{r}
# run one-sided t-test with null hypothesis of mean as 60, alternative less than 60
results <- my_t.test(my_gapminder$lifeExp, "less", 60)
results
```

Using a p-value cut-off at $\alpha = 0.05$, we can see that a p-value of 
`r results[[4]]` implies that we can reject the null hypothesis. 
In other words, we have sufficient evidence to reject the null hypothesis that
the mean life expectancy is 60 years in support of the alternative hypothesis
that the mean life expectancy is less than 60 years.

### one-sided test, greater

We run a t-test with the null hypothesis that the mean life expectancy is 60 years,
with the alternative hypothesis being that the life expectancy is greater than 60 years.

```{r}
# run one-sided t-test with null hypothesis of mean as 60, alternative greater than 60
results <- my_t.test(my_gapminder$lifeExp, "greater", 60)
results
```

Using a p-value cut-off at $\alpha = 0.05$, we can see that a p-value of 
`r results[[4]]` implies that we cannot reject the null hypothesis. 
In other words, we do not have sufficient evidence to reject the null hypothesis that
the mean life expectancy is 60 years in support of the alternative hypothesis
that the mean life expectancy is greater than 60 years.

# ```my_lm``` Tutorial
