---
title: "Project 3: STAT302PACKAGE Tutorial"
author: "David Zeng"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{STAT302PACKAGE Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(STAT302PACKAGE)
```

# Introduction

The STAT302PACKAGE contains four main functions (in addition to some 
other minor ones):

* ```my_t.test```, a function which computes a one sample t-test 
* ```my_lm```, a function which fits a linear model 
* ```my_knn_Cv```, a function which trains a n-nearest neighbors model using 
k-fold cross-validation
* ```my_rf_cv``, a function which trains a random forest model using k-fold 
cross-validation

To install and use this package, run the following commands in console:

```{r, eval=FALSE}
# install and load package
devtools::install_github("dzeng8/STAT302PACKAGE")
library(STAT302PACKAGE)
```

# ```my_t.test``` Tutorial

